---
title: "Scraping data from Twitter's REST API"
author: "Pablo Barbera"
date: "December 18, 2017"
output: html_document
---

We'll now turn to a different type of Twitter data -- static data, either recent tweets or user-level information. This type of data can be retrieved with Twitter's REST API. We will use the `netdemR` package here -- this is a package that we created in the NetDem lab at USC to facilitate the collection and analysis of Twitter data.

### Searching recent tweets

It is possible to download recent tweets, but only up those less than 7 days old, and in some cases not all of them.

```{r}
library(netdemR)
library(streamR)

searchTweets(q=c("doug jones", "roy moore"), 
  filename="../data/alabama-tweets.json",
  n=1000, until="2017-12-15", 
  oauth_folder="../credentials")

tweets <- parseTweets("../data/alabama-tweets.json")
```

What are the most popular hashtags?
```{r}
library(stringr)
ht <- str_extract_all(tweets$text, "#(\\d|\\w)+")
ht <- unlist(ht)
head(sort(table(ht), decreasing = TRUE))
```

You can check the documentation about the options for string search [here](https://dev.twitter.com/rest/public/search).

### Extracting users' profile information

This is how you would extract information from user profiles:

```{r}
wh <- c("realDonaldTrump", "POTUS", "VP", "FLOTUS")
users <- getUsersBatch(screen_names=wh,
                       oauth_folder="../credentials")
str(users)
```

Which of these has the most followers?
```{r}
users[which.max(users$followers_count),]
users$screen_name[which.max(users$followers_count)]
```

Download up to 3,200 recent tweets from a Twitter account:
```{r}
getTimeline(filename="../data/realDonaldTrump.json", screen_name="realDonaldTrump", 
    n=1000, oauth_folder="../credentials")
```

What are the most common hashtags?
```{r}
tweets <- parseTweets("../data/realDonaldTrump.json")
ht <- str_extract_all(tweets$text, "#(\\d|\\w)+")
ht <- unlist(ht)
head(sort(table(ht), decreasing = TRUE))
```

### Building friend and follower networks

Download friends and followers:
```{r}
followers <- getFollowers("BJPolS", 
    oauth_folder="../credentials")
friends <- getFriends("BJPolS", 
    oauth_folder="../credentials")
```

What are the most common words that friends of the BJPS Twitter account use to describe themselves on Twitter?

```{r, fig.height=6, fig.width=6}
# extract profile descriptions
users <- getUsersBatch(ids=friends, oauth_folder="../credentials")
# create table with frequency of word use
library(quanteda)
tw <- corpus(users$description[users$description!=""])
dfm <- dfm(tw, remove=c(stopwords("english"), stopwords("spanish"),
                                 "t.co", "https", "rt", "rts", "http"),
           remove_punct=TRUE)
# create wordcloud
par(mar=c(0,0,0,0))
textplot_wordcloud(dfm, rot.per=0, scale=c(3, .50), max.words=100)

```

### Other types of data

The REST API offers also a long list of other endpoints that could be of use at some point, depending on your research interests.

1) You can search users related to specific keywords:

```{r}
users <- searchUsers(q="political science", count=100, oauth_folder="../credentials")
users$screen_name[1:10]
```

2) If you know the ID of the tweets, you can download it directly from the API. This is useful because tweets cannot be redistributed as part of the replication materials of a published paper, but the list of tweet IDs can be shared:

```{r}
# Downloading tweets when you know the ID
getStatuses(ids=c("474134260149157888", "266038556504494082"),
            filename="../data/old-tweets.json",
            oauth_folder="../credentials")
parseTweets("../data/old-tweets.json")
```

3) Lists of Twitter users, compiled by other users, are also accessible through the API.

```{r}
# download user information from a list
MCs <- getList(list_name="new-members-of-congress", 
               screen_name="cspan", oauth_folder="../credentials")
head(MCs)
```

This is also useful if e.g. you're interested in compiling lists of journalists, because media outlets offer these lists in their profiles.

4) List of users who retweeted a particular tweet -- unfortunately, it's limited to only 100 most recent retweets.

```{r}
# Download list of users who retweeted a tweet (unfortunately, only up to 100)
rts <- getRetweets(id='942123433873281024', oauth_folder="../credentials")
# https://twitter.com/realDonaldTrump/status/942123433873281024
users <- getUsersBatch(ids=rts, oauth_folder="../credentials")
# create table with frequency of word use
library(quanteda)
tw <- corpus(users$description[users$description!=""])
dfm <- dfm(tw, remove=c(stopwords("english"), stopwords("spanish"),
                                 "t.co", "https", "rt", "rts", "http"),
           remove_punct = TRUE)
# create wordcloud
par(mar=c(0,0,0,0))
textplot_wordcloud(dfm, rot.per=0, scale=c(5, .50), max.words=100)
```

5) And one final function to convert dates in their internal Twitter format to another format we could work with in R:

```{r}
# format Twitter dates to facilitate analysis
tweets <- parseTweets("../data/realDonaldTrump.json")
tweets$date <- formatTwDate(tweets$created_at, format="date")
hist(tweets$date, breaks="month")
```

Now time for our second challenge of the day!

